A Backsolving Technique Using Dynamic Taint Analysis

### 2013-12-03 docs/paper/fsuthesis/thesis-template/paper.tex
docs/paper/fsuthesis/thesis-template/paper.tex has my paper

%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
%               sigplanconf-guide.pdf
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------

\documentclass[10pt, twocolumn, times, nocopyrightspace, preprint]{sigplanconf}

% The following \documentclass options may be useful:

% preprint      Remove this option only once the paper is in final form.
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{amsmath}

@@

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country} 
\copyrightyear{20yy} 
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
\doi{nnnnnnn.nnnnnnn}

% Uncomment one of the following two, if you are not going for the 
% traditional copyright transfer agreement.

%\exclusivelicense                % ACM gets exclusive license to publish, 
                                  % you retain copyright

%\permissiontopublish             % ACM gets nonexclusive license to publish
                                  % (paid open-access papers, 
                                  % short abstracts)

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Pseudo-Symbolic Execution (Or Minimialistic Constraint Solving) Using Dynamic Taint Analysis and Backsolving}
%\subtitle{Subtitle Text, if any}

\authorinfo{Clark Wood}
           {Florida State University}
           {clark.w.wood@gmail.com}
%\authorinfo{Name2\and Name3}
%           {Affiliation2/3}
%           {Email2/3}

\maketitle

%\begin{abstract}
%This is the text of the abstract.
%\end{abstract}

\category{CR-number}{subcategory}{third-level}

% general terms are not compulsory anymore, 
% you may leave them out
%\terms
%term1, term2

%\keywords
%keyword1, keyword2

@@

Finding exploitable bugs in binaries is a difficult and lucrative endeavour. Fuzzing, or the feeding of random input into processes to try and detect vulnerable conditions [CITE], is one technique used by researchers and analysts. Applications tested for vulnerabilities are often black boxes because source code is unavailable. This piles difficulties onto an already complex problem, of which much of current research attempts to solve or reduce [CITES]. Among these issues are i) efficient and thorough dynamic taint analysis [CITE], ii) the path explosion problem [U], and iii) the code coverage problem [CITE?]. Dynamic taint analysis is difficult because taint can propagate directly via assignment, or indirectly via affecting control flows which affect assignment [D], which make proper implementation tricky. The path explosion problem arises from the nature of assembly languages. Jump instructions, which may make up as much as X\% of machine instructions in programs [CITE], each multiply the total number of existing paths in a binary by two. Considering the size of binaries, this quickly leads to an un-tenably large set of possible paths to explore via conventional techniques like symbolic execution. And even if it is possible to explore all paths in a binary, analysts still need sample input which can trigger the vulnerable condition, perhaps proving that the binary is exploitable along the way. Such automated exploit generation is the subject of much research [H, I, CITES].

This work aims to reduce the severity of the latter two problems while being wary of the first during implementation. Instead of trying to attain more reasonable performance by using symbolic execution less or more wisely [J], I focus instead on approximating symbolic execution by using dynamic taint analysis to drive mutation. This is possible by tracking taint at the instruction level, then backsolving from the jump to the source of taint to come up with smarter ways to mutate user input and create new test cases. For instance, in the below example:

\begin{verbatim}
mv eax, userInput1
mv ebx, userInput2
cmp eax, ebx
jge 0xDEADBEEF
\end{verbatim}

If we already had a test case where userInput1 equalled 10 and userInput2 equalled 5, this test case would jump to 0xDEADBEEF. To generate a test case where we would not jump, we take the complement of the jge instruction's comparison function, which would be "less than". We can find the last instruction that modified the Z flag, which is how jge decides to jump or not. We see that this compare instruction was tained by the eax and ebx registers, which were in turn tainted by userInputs 1 and 2. Since we can impact both values in the comparison, it is easy to modify user input smartly to alter the value of the Z flag and jump during the next execution. Perhaps we could switch the values, or auto-increment/decrement one until we satisfy the needed comparison, or mutate values randomly.

@@

There are a finite number of ways to compare data in x86, so I can create cases for all possible ways jumps are decided. I only want to prove this is possible however, so I will only implement a subset and instead comment on how to implement the other comparisons.

@@

I assume ASLR is disabled for the binary to be fuzzed. This allows the program to easily track and compare execution paths by creating a list of jump locations. Since the world is moving more and more toward 64-bit architectures, I plan to implement the technique for x86\_64.

@@

I am leveraging Intel PIN to perform dynamic binary instrumentation. I plan to test first on dummy programs with many paths that are easy to backsolve, similar to the example above. A C program which uses sequential switch statements to check user input against a secret, when compiled with tcc, which performs very little optimization, should create an appropriate binary. I then plan to test against S2E and Peach Fuzzer. I already have this program written and have verified using IDA Pro that there are a many, many paths to try and find.

@@

\section{Research}

@@

Dynamic taint analysis (DTA) is the process of locating sources of taint and following their propagation through a binary to data sinks. Since, for exploitation, bugs must be triggered by user input, DTA usually identifies all user input (files, command line arguments, UI interactions, ...) as tainted, and marks it accordingly.

@@

I plan to implement dynamic taint analysis for files as input sources. I am basing my DTA heavily on Jonathan Salwan\'s work [N]. I've been in contact with him and received permission to use his sample code. I can watch for all reads after a file is opened, then tag user input at the byte level and follow it as it moves throughout the binary.

I will need to implement a system which keeps track of how taint propagates at the instruction level. A directed graph structure, with initial nodes representing the bytes of user input and edges pointing to subsequent instructions which propagate taint, along with some metadata to make backsolving easier, seems like a reasonable data structure.

@@

\section{Setbacks}

@@

\begin{itemize}
	\item [A] Bekrar, Sofia, et al. "A taint based approach for smart fuzzing." Software Testing, Verification and Validation (ICST), 2012 IEEE Fifth International Conference on. IEEE, 2012.
	\item [B] DeMott, J., Enbody, R., and Punch, W. "Revolutionizing the Field of Grey-box Attack Surface Testing with
Evolutionary Fuzzing", BlackHat and Defcon 2007.
	\item [C] Schwartz, E.J., Avgerinos, T., Brumley, D. "All You Ever Wanted to Know About Dynamic Taint Analysis and Forward Symbolic Execution (but might have been afraid to ask)." 2010 IEEE Symposium on Security and Privacy.
	\item [D] Clause, J. Li, W., Orso, A. "Dytan: a generic dynamic taint analysis framework". 2007 Int'l symposium on Software testing and analysis. ACM, 2007.
	\item [E] ... "Beyond Instruction Level Taint Propagation".
	\item [F] Yamaguchi, Fabian, Felix Lindner, and Konrad Rieck. "Vulnerability extrapolation: assisted discovery of vulnerabilities using machine learning." Proceedings of the 5th USENIX conference on Offensive technologies. USENIX Association, 2011.
	\item [G] Bao, Tao, et al. "Strict control dependence and its effect on dynamic information flow analyses." Proceedings of the 19th international symposium on Software testing and analysis. ACM, 2010.
	\item [H] Cha, Sang Kil, et al. "Unleashing mayhem on binary code." Security and Privacy (SP), 2012 IEEE Symposium on. IEEE, 2012.
	\item [I] Avgerinos, Thanassis, et al. "AEG: Automatic Exploit Generation." NDSS. Vol. 11. 2011.
	\item [J] Chipounov, Vitaly, Volodymyr Kuznetsov, and George Candea. "S2E: A platform for in-vivo multi-path analysis of software systems." ACM SIGARCH Computer Architecture News 39.1 (2011): 265-278.
	\item [K] Bellard, Fabrice. "QEMU, a Fast and Portable Dynamic Translator." USENIX Annual Technical Conference, FREENIX Track. 2005.
	\item [L] Guillon, Christophe. "Program Instrumentation with QEMU." 1st International QEMU Users’ Forum. 2011.
	\item [M] Luk, Chi-Keung, et al. "Pin: building customized program analysis tools with dynamic instrumentation." Acm Sigplan Notices. Vol. 40. No. 6. ACM, 2005.
	\item [N] Salwan, Jonathan. Shell-storm.org %\url{http://shell-storm.org/blog/Taint-analysis-and-pattern-matching-with-Pin/}
	\item [O] King, James C. "Symbolic execution and program testing." Communications of the ACM 19.7 (1976): 385-394.
	\item [P] De Moura, Leonardo, and Nikolaj Bjørner. "Z3: An efficient SMT solver." Tools and Algorithms for the Construction and Analysis of Systems. Springer Berlin Heidelberg, 2008. 337-340.
	\item [Q] Godefroid, Patrice, Michael Y. Levin, and David Molnar. "Sage: Whitebox fuzzing for security testing." Queue 10.1 (2012): 20.
	\item [R] Ma, Kin-Keung, et al. "Directed symbolic execution." Static Analysis. Springer Berlin Heidelberg, 2011. 95-111.
	\item [S] Boyer, Robert S., Bernard Elspas, and Karl N. Levitt. "SELECT—a formal system for testing and debugging programs by symbolic execution." ACM SigPlan Notices. Vol. 10. No. 6. ACM, 1975.
	\item [T] Wang, Tielei, et al. "Checksum-aware fuzzing combined with dynamic taint analysis and symbolic execution." ACM Transactions on Information and System Security (TISSEC) 14.2 (2011): 15.
	\item [U] Cadar, Cristian, and Koushik Sen. "Symbolic execution for software testing: three decades later." Communications of the ACM 56.2 (2013): 82-90.
	\item [V] Cadar, Cristian, Daniel Dunbar, and Dawson R. Engler. "KLEE: Unassisted and Automatic Generation of High-Coverage Tests for Complex Systems Programs." OSDI. Vol. 8. 2008.
	\item [W] McCamant, Stephen, et al. Transformation-aware symbolic execution for system test generation. Tech. Rep. UCB/EECS-2013-125, University of California, Berkeley (Jun 2013), 2013.
	\item [X] Majumdar, Rupak, and Koushik Sen. "Hybrid concolic testing." Software Engineering, 2007. ICSE 2007. 29th International Conference on. IEEE, 2007.
	\item [Y] Sen, Koushik, Darko Marinov, and Gul Agha. CUTE: a concolic unit testing engine for C. Vol. 30. No. 5. ACM, 2005.

@@

%\appendix
%\section{Appendix Title}

%This is the text of the appendix, if you need one.

%\acks

%Acknowledgments, if needed.

% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrvnat}

% The bibliography should be embedded for final submission.

%\begin{thebibliography}{}
%\softraggedright

%\bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
%P. Q. Smith, and X. Y. Jones. ...reference text...

%\end{thebibliography}

@@

%                       Revision History
%                       -------- -------
%  Date         Person  Ver.    Change
%  ----         ------  ----    ------

%  2013.06.29   TU      0.1--4  comments on permission/copyright notices

>>>>

% This is a "bare-bones" thesis template file.  For examples of how to
% use a few more LaTeX features, look in the 'sample' folder.  Read
% the User Guide for documentation of the 'fsuthesis' class features.

\documentclass[11pt,expanded,copyright]{fsuthesis}

% Additional packages may be loaded here.

\title{A Backsolving Technique Using Dynamic Taint Analysis}
\author{Clark Wood}
\college{College of Arty Science}
\department{Department of Computer Science} % Delete if no department
\manuscripttype{Thesis}              % [Thesis, Dissertation, Treatise]
\degree{Masters of Science}                 
\semester{Spring}                            % [Spring, Summer, Fall]
\degreeyear{2014}
\defensedate{Spring 2014}

%\subject{My Topic}
%\keywords{keyterm1; keyterm2; keyterm3; ...}

%\committeeperson{Faux Causson Yorverk}{Professor Directing Thesis}
%\committeeperson{Verda Boizaar}{University Representative}
%\committeeperson{Beauxeau D'Claune}{Committee Member}
%\committeeperson{Arlip Zarseeld}{Committee Member}

\clubpenalty=9999
\widowpenalty=9999

@@

\frontmatter
\maketitle
\makecommitteepage

%\begin{dedication}
%\end{dedication}

%\begin{acknowledgments}
%\end{acknowledgments}

\tableofcontents
%\listoftables
%\listoffigures
%\listofmusex

%\begin{listofsymbols}
%\end{listofsymbols}

%\begin{listofabbrevs}
%\end{listofabbrevs}

%\begin{abstract}
%This is my abstract.
%\end{abstract}

\mainmatter

%\input chapter1

@@

Finding exploitable bugs in binaries is a difficult and lucrative endeavour. Fuzzing, or the feeding of controlled input into processes to try and detect vulnerable conditions [CITE], is one technique used by researchers and analysts. Applications tested for vulnerabilities are often black boxes because source code is unavailable. This piles difficulties onto an already complex problem, of which much of current research attempts to solve or reduce [CITES]. Among these issues are i) efficient and thorough dynamic taint analysis [CITE], ii) the path explosion problem [U], and iii) the code coverage problem [CITE?]. Dynamic taint analysis is difficult because taint can propagate directly via assignment, or indirectly via affecting control flows which affect assignment [D], which make proper implementation tricky. The path explosion problem arises from the nature of assembly languages. Jump instructions, which may make up as much as X\% of machine instructions in programs [CITE], each multiply the total number of existing paths in a binary by two. Considering the size of binaries, this quickly leads to an un-tenably large set of possible paths to explore via conventional techniques like symbolic execution. And even if it is possible to explore all paths in a binary, analysts still need sample input which can trigger the vulnerable condition, perhaps proving that the binary is exploitable along the way. Such automated exploit generation is the subject of much research [H, I, CITES].

This work aims to reduce the severity of the latter two problems while being wary of the first during implementation. Instead of trying to attain more reasonable performance by using symbolic execution less or more wisely [J], I focus instead on approximating symbolic execution by using dynamic taint analysis to drive mutation. This is possible by tracking taint at the instruction level, then backsolving from the jump to the source of taint to come up with smarter ways to mutate user input and create new test cases. For instance, in the figure 1 example:

\begin{verbatim}
figure 1:
call read
mov eax, [ebp - val]
inc eax
mul eax, 5
cmp eax, 42
jnz 0x8BADMEAT
\end{verbatim}

If we already had a test case where the value read onto the stack, after the transformations, did not equal 42, then the cmp instruction would not set the zero flag, and we would jump to the memory location 0xBADMEAT. With dynamic taint analysis in place, we see taint introduced with the read syscall, then moved into the eax register (which is subsequently tainted). After some number of transformations, perhaps zero, to the value in the eax register, we see a compare instruction, which checks eax's value against an immediate value. Then we can keep track of specific taint sources, and can record any movements or transformations, we have all the necessary information to "solve" the branch. Once we reach the tainted branch, we know that user input can affect whether we jump or not, and consequently we can backsolve to find out how to affect the branch instruction. In particular, we know the conditional jump instruction jnz depends on the Z flag in the EFLAGS register. We record whatever instruction last modified each flag in the Z flags register, so we handle the compare instruction next. Because we have kept track of taint propagation, we know we can affect the value of eax, but not 42, since it is an immediate value, so we know we need to make the value in eax equal 42. We follow the instructions that have modified the value of eax and use any appropriate operands involved with the inverse operation of the one originally performed on the value to reverse the transformations to eax's value. Finally, we notice eax originally received taint when a value was moved from memory, which was read into memory using the read syscall.

@@

Because there are a finite number of conditional jumps, all of which depend on different flags or combinations of flags in the EFLAGS register, and a finite number of instructions that modify the EFLAGS register in x86, I can create cases for all possible ways jumps are decided. I only want to prove this is possible however, so I will only implement a subset and instead comment on how to implement the other comparisons.

\section{Related Work}

The major impetus for this work is outlined in "A Taint Based Approach for Smart Fuzzing". This paper proposes a fuzzing tool architecture, involving vulnerability detection, followed by taint analysis based upon found vulnerabilities, which drives the generation of intelligent tests. As tests are run they are checked to ensure they provide adequate code coverage, and results are carefully monitored for interesting situations like crashes [A]. The paper also serves as an excellent explanation of the current state and future of fuzzing. In particular, they suggest combining taint analysis with backward slicing. The technique discussed in the paper boils down to an attempt to automatically solve slices like those introduced in [AA], which have been derived from taint analysis as much as possible.

@@

I assume ASLR is disabled for the binary to be fuzzed. This allows the program to easily track and compare execution paths by creating a list of jump locations. Since the world is moving more and more toward 64-bit architectures, I plan to implement the technique for 32-bit x86 assembly. This is because x86 is widespread, and vulnerabilities resulting from unsafe C code should persist in binaries, regardless of both 32 vs. 64-bit and the flavor of architecture (ARM, x86, PowerPC), since compilers try to adhere to the C standard.

@@

I am leveraging Intel PIN to perform dynamic binary instrumentation. I plan to test first on dummy programs with many paths that are easy to backsolve, similar to the example above. A C program which uses sequential switch statements to check user input against a secret, when compiled with tcc, which performs very little optimization, should create an appropriate binary. I then plan to test against S2E and Peach Fuzzer [Z]. I already have this program written and have verified using IDA Pro that there are a many, many paths to try and find.

@@

\section{Techniques}

Several valuable techniques to improve fuzzers' code coverage have evolved over the years. Two, in particular, inspire this work directly: Dynamic Taint Analysis and Concolic Execution. Dynamic Taint Analysis, or DTA, provides valuable information for guided fuzzing. Concolic execution, which has spawned out of the desire to use symbolic execution more wisely due to performance constraints, aims to ameliorate the path explosion problem. These each come with their own drawbacks, a primary one of which is the lack of automated instrumentation. Historically, symbolic execution required a programmer to directly instrument source code, but much research is now focused on automating as much as possible.

@@

DTA is the process of locating sources of taint and following their propagation through a binary to data sinks. Since, for exploitation, bugs must be triggered by user input, DTA usually identifies all user input (files, command line arguments, UI interactions, ...) as tainted, and marks it accordingly.

@@

I implement dynamic taint analysis for files as input sources. I am basing my DTA heavily on Jonathan Salwan\'s work [N]. Currently I instrument the read syscall, watching for all reads executed, then tag user input at the byte level and follow it as it moves throughout the binary. I instrument binaries at the assembly instruction level using Intel Pin. Each instruction is handled differently based upon the following criteria:

\begin{itemize}
	\item Is the instruction a conditional jump?
	\item Is the first operand a memory location or a register?
	\item Is the first operand read or written?
	\item Is the first (second if applicable) operand tainted?
\end{itemize}

Operations involving an untainted first operand and a tainted second operand generally spread taint, while operations with a tainted first operand and an untainted second operand remove taint. Tainted branches trigger the backsolving technique. Various precautions, such as removing taint for an instruction like "xor eax, eax", must be taken. Lossy operations, like shifts, must be handled carefully to be backsolve-able.

\subsection{Testing Harness}

Fuzzers require intuitive but powerful test harnesses, so that users can easily control how a binary is fuzzed and understand the results of fuzz testing. Currently, I record input, output, and taint for each test executed, using an auto-incrementing id and separating results at the directory level. I also keep track of the mutations which have already been explored. Users can grep for interesting conditions like segfaults or unexpected output, then cross reference the test number with the observed taint flow and the particular mutation tested.

Professional fuzzing products, like Peach Fuzzer

\subsection{Methodology}

@@

\subsection{Evaluation}

%
%Currently have results for crackmes, give idea of my fuzzer vs. %peach fuzz, and where i plan to go.

\section{Future Work}

% Was setbacks, modify and remove heavily

@@

\begin{itemize}
	\item [A] Bekrar, Sofia, et al. "A taint based approach for smart fuzzing." Software Testing, Verification and Validation (ICST), 2012 IEEE Fifth International Conference on. IEEE, 2012.
	\item [B] DeMott, J., Enbody, R., and Punch, W. "Revolutionizing the Field of Grey-box Attack Surface Testing with
Evolutionary Fuzzing", BlackHat and Defcon 2007.
	\item [C] Schwartz, E.J., Avgerinos, T., Brumley, D. "All You Ever Wanted to Know About Dynamic Taint Analysis and Forward Symbolic Execution (but might have been afraid to ask)." 2010 IEEE Symposium on Security and Privacy.
	\item [D] Clause, J. Li, W., Orso, A. "Dytan: a generic dynamic taint analysis framework". 2007 Int'l symposium on Software testing and analysis. ACM, 2007.
	\item [E] ... "Beyond Instruction Level Taint Propagation".
	\item [F] Yamaguchi, Fabian, Felix Lindner, and Konrad Rieck. "Vulnerability extrapolation: assisted discovery of vulnerabilities using machine learning." Proceedings of the 5th USENIX conference on Offensive technologies. USENIX Association, 2011.
	\item [G] Bao, Tao, et al. "Strict control dependence and its effect on dynamic information flow analyses." Proceedings of the 19th international symposium on Software testing and analysis. ACM, 2010.
	\item [H] Cha, Sang Kil, et al. "Unleashing mayhem on binary code." Security and Privacy (SP), 2012 IEEE Symposium on. IEEE, 2012.
	\item [I] Avgerinos, Thanassis, et al. "AEG: Automatic Exploit Generation." NDSS. Vol. 11. 2011.
	\item [J] Chipounov, Vitaly, Volodymyr Kuznetsov, and George Candea. "S2E: A platform for in-vivo multi-path analysis of software systems." ACM SIGARCH Computer Architecture News 39.1 (2011): 265-278.
	\item [K] Bellard, Fabrice. "QEMU, a Fast and Portable Dynamic Translator." USENIX Annual Technical Conference, FREENIX Track. 2005.
	\item [L] Guillon, Christophe. "Program Instrumentation with QEMU." 1st International QEMU Users’ Forum. 2011.
	\item [M] Luk, Chi-Keung, et al. "Pin: building customized program analysis tools with dynamic instrumentation." Acm Sigplan Notices. Vol. 40. No. 6. ACM, 2005.
	\item [N] Salwan, Jonathan. Shell-storm.org %\url{http://shell-storm.org/blog/Taint-analysis-and-pattern-matching-with-Pin/}
	\item [O] King, James C. "Symbolic execution and program testing." Communications of the ACM 19.7 (1976): 385-394.
	\item [P] De Moura, Leonardo, and Nikolaj Bjørner. "Z3: An efficient SMT solver." Tools and Algorithms for the Construction and Analysis of Systems. Springer Berlin Heidelberg, 2008. 337-340.
	\item [Q] Godefroid, Patrice, Michael Y. Levin, and David Molnar. "Sage: Whitebox fuzzing for security testing." Queue 10.1 (2012): 20.
	\item [R] Ma, Kin-Keung, et al. "Directed symbolic execution." Static Analysis. Springer Berlin Heidelberg, 2011. 95-111.
	\item [S] Boyer, Robert S., Bernard Elspas, and Karl N. Levitt. "SELECT—a formal system for testing and debugging programs by symbolic execution." ACM SigPlan Notices. Vol. 10. No. 6. ACM, 1975.
	\item [T] Wang, Tielei, et al. "Checksum-aware fuzzing combined with dynamic taint analysis and symbolic execution." ACM Transactions on Information and System Security (TISSEC) 14.2 (2011): 15.
	\item [U] Cadar, Cristian, and Koushik Sen. "Symbolic execution for software testing: three decades later." Communications of the ACM 56.2 (2013): 82-90.
	\item [V] Cadar, Cristian, Daniel Dunbar, and Dawson R. Engler. "KLEE: Unassisted and Automatic Generation of High-Coverage Tests for Complex Systems Programs." OSDI. Vol. 8. 2008.
	\item [W] McCamant, Stephen, et al. Transformation-aware symbolic execution for system test generation. Tech. Rep. UCB/EECS-2013-125, University of California, Berkeley (Jun 2013), 2013.
	\item [X] Majumdar, Rupak, and Koushik Sen. "Hybrid concolic testing." Software Engineering, 2007. ICSE 2007. 29th International Conference on. IEEE, 2007.
	\item [Y] Sen, Koushik, Darko Marinov, and Gul Agha. CUTE: a concolic unit testing engine for C. Vol. 30. No. 5. ACM, 2005.
	\item [Z] Peach Fuzzer. url{http://peachfuzzer.com/}
	\item [AA] Zhang, Xiangyu, Rajiv Gupta, and Youtao Zhang. "Precise dynamic slicing algorithms." Software Engineering, 2003. Proceedings. 25th International Conference on. IEEE, 2003.

@@

%\input chapter2
%\input chapter3

%\appendix
%\input appendix1

% You have your choice of bibliography sections, either
% hand-crafted or BibTeX.

% This is the "hand-crafted" bibliography/references section:
%\begin{references}
%Mybib, Sample. \textit{An Example of a Bibliographic Entry
% Created Manually}. Tallahassee, Florida: Fornish and Frak, 2010.

%Smith, Marigold. \textit{Lots and Lots of Bibliographic Entries
% and How to Display Them}. Tallahassee, Florida: Gibson and %Goulash, 2010.
%\end{references}

% Or use the BibTeX bibliography/references section below.  View the
% file 'myrefs.bib' to get a feel for what these entries may look
% like.  See the document in the 'sample' folder for more citation and
% BibTeX examples.

% \bibliographystyle{plain}
% \bibliography{myrefs}

\begin{biosketch}
This is my biography.
\end{biosketch}

@@

### 2013-12-03 docs/paper/sigplanconf-template.tex
docs/paper/fsuthesis/thesis-template/paper.tex has my paper

\conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country} 
\copyrightyear{20yy} 
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
\doi{nnnnnnn.nnnnnnn}

@@

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Pseudo-Symbolic Execution (Or Minimialistic Constraint Solving) Using Dynamic Taint Analysis and Backsolving}
%\subtitle{Subtitle Text, if any}

@@

@@

Finding exploitable bugs in binaries is a difficult and lucrative endeavour. Fuzzing, or the feeding of random input into processes to try and detect vulnerable conditions [CITE], is one technique used by researchers and analysts. Applications tested for vulnerabilities are often black boxes because source code is unavailable. This piles difficulties onto an already complex problem, of which much of current research attempts to solve or reduce [CITES]. Among these issues are i) efficient and thorough dynamic taint analysis [CITE], ii) the path explosion problem [U], and iii) the code coverage problem [CITE?]. Dynamic taint analysis is difficult because taint can propagate directly via assignment, or indirectly via affecting control flows which affect assignment [D], which make proper implementation tricky. The path explosion problem arises from the nature of assembly languages. Jump instructions, which may make up as much as X\% of machine instructions in programs [CITE], each multiply the total number of existing paths in a binary by two. Considering the size of binaries, this quickly leads to an un-tenably large set of possible paths to explore via conventional techniques like symbolic execution. And even if it is possible to explore all paths in a binary, analysts still need sample input which can trigger the vulnerable condition, perhaps proving that the binary is exploitable along the way. Such automated exploit generation is the subject of much research [H, I, CITES].

This work aims to reduce the severity of the latter two problems while being wary of the first during implementation. Instead of trying to attain more reasonable performance by using symbolic execution less or more wisely [J], I focus instead on approximating symbolic execution by using dynamic taint analysis to drive mutation. This is possible by tracking taint at the instruction level, then backsolving from the jump to the source of taint to come up with smarter ways to mutate user input and create new test cases. For instance, in the below example:

\begin{verbatim}
mv eax, userInput1
mv ebx, userInput2
cmp eax, ebx
jge 0xDEADBEEF
\end{verbatim}

If we already had a test case where userInput1 equalled 10 and userInput2 equalled 5, this test case would jump to 0xDEADBEEF. To generate a test case where we would not jump, we take the complement of the jge instruction's comparison function, which would be "less than". We can find the last instruction that modified the Z flag, which is how jge decides to jump or not. We see that this compare instruction was tained by the eax and ebx registers, which were in turn tainted by userInputs 1 and 2. Since we can impact both values in the comparison, it is easy to modify user input smartly to alter the value of the Z flag and jump during the next execution. Perhaps we could switch the values, or auto-increment/decrement one until we satisfy the needed comparison, or mutate values randomly.

>>>>

%\conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country} 
%\copyrightyear{20yy} 
%\copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
%\doi{nnnnnnn.nnnnnnn}

@@

%\titlebanner{banner above paper title}        % These are ignored unless
%\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{A Backsolving Technique Using Dynamic Taint Analysis}
%\subtitle{Subtitle Text, if any}

@@

\section{Abstract}

@@

Finding exploitable bugs in binaries is a difficult and lucrative endeavour. Fuzzing, or the feeding of controlled input into processes to try and detect vulnerable conditions [CITE], is one technique used by researchers and analysts. Applications tested for vulnerabilities are often black boxes because source code is unavailable. This piles difficulties onto an already complex problem, of which much of current research attempts to solve or reduce [CITES]. Among these issues are i) efficient and thorough dynamic taint analysis [CITE], ii) the path explosion problem [U], and iii) the code coverage problem [CITE?]. Dynamic taint analysis is difficult because taint can propagate directly via assignment, or indirectly via affecting control flows which affect assignment [D], which make proper implementation tricky. The path explosion problem arises from the nature of assembly languages. Jump instructions, which may make up as much as X\% of machine instructions in programs [CITE], each multiply the total number of existing paths in a binary by two. Considering the size of binaries, this quickly leads to an un-tenably large set of possible paths to explore via conventional techniques like symbolic execution. And even if it is possible to explore all paths in a binary, analysts still need sample input which can trigger the vulnerable condition, perhaps proving that the binary is exploitable along the way. Such automated exploit generation is the subject of much research [H, I, CITES].

This work aims to reduce the severity of the latter two problems while being wary of the first during implementation. Instead of trying to attain more reasonable performance by using symbolic execution less or more wisely [J], I focus instead on approximating symbolic execution by using dynamic taint analysis to drive mutation. This is possible by tracking taint at the instruction level, then backsolving from the jump to the source of taint to come up with smarter ways to mutate user input and create new test cases. For instance, in the figure 1 example:
\\
\\

\begin{verbatim}
figure 1:
call read
mov eax, [ebp - val]
inc eax
mul eax, 5
cmp eax, 42
jnz 0x8BADMEAT
\end{verbatim}

If we already had a test case where the value read onto the stack, after the transformations, did not equal 42, then the cmp instruction would not set the zero flag, and we would jump to the memory location 0xBADMEAT. With dynamic taint analysis in place, we see taint introduced with the read syscall, then moved into the eax register (which is subsequently tainted). After some number of transformations, perhaps zero, to the value in the eax register, we see a compare instruction, which checks eax's value against an immediate value. Then we can keep track of specific taint sources, and can record any movements or transformations, we have all the necessary information to "solve" the branch. Once we reach the tainted branch, we know that user input can affect whether we jump or not, and consequently we can backsolve to find out how to affect the branch instruction. In particular, we know the conditional jump instruction jnz depends on the Z flag in the EFLAGS register. We record whatever instruction last modified each flag in the Z flags register, so we handle the compare instruction next. Because we have kept track of taint propagation, we know we can affect the value of eax, but not 42, since it is an immediate value, so we know we need to make the value in eax equal 42. We follow the instructions that have modified the value of eax and use any appropriate operands involved with the inverse operation of the one originally performed on the value to reverse the transformations to eax's value. Finally, we notice eax originally received taint when a value was moved from memory, which was read into memory using the read syscall.
